---
title: "Practical Machine Learning Project"
author: "Matías"
date: "21 de octubre de 2015"
output: html_document
---


#Loading the Data
```{r}
library(caret)
set.seed(4433)
traindata <- read.csv("pml-training.csv", header = T, sep = ",")
```

#Cleaning Data

##Removing variables with near zero varibilty 
```{r}
dim(traindata)
nzv <- nearZeroVar(traindata)
filtereddata <- traindata[, -nzv]

dim(filtereddata)
```

##Removing variables with more than 14000 NA's
```{r}
c<-c()
for (i in  1:100) {
       if(sum(is.na(filtereddata[, i]))>14000) 
          c <- c(c,i)
       }
new_data <- filtereddata[,-c]
```

##Removing id, name and time stamp; this variables are useless
```{r}
new_data <- new_data[, -c(1,2)]
new_data <- new_data[,-3]
```
#Creating train and test partitions.
###I create a training and a testing data set from the pml-training file.

```{r}
inTrain <- createDataPartition(y=new_data$classe, p=0.75, list = FALSE)
training <- new_data[inTrain,]
testing <- new_data[-inTrain,]
```

###There are some highly correlated variables, but since I'm going to use cart and random forest I choose to leave them.
```{r}
cor_mat <- cor(subset(new_data, select = -classe))
highlycorrelated <- findCorrelation(cor_mat, cutoff = 0.75)
names(new_data)[highlycorrelated]
highlycorrelated 
```
#Training the model.
```{r}
library(rpart)
modelFit1 <- rpart(classe ~ ., data=training, method="class")
```

```{r}
library(rattle)
fancyRpartPlot(modelFit1)
```
#Evaluating the model.
```{r}
predictions1 <- predict(modelFit1, testing, type = "class")
confusionMatrix(predictions1, testing$classe)
```
###Doing the confusion Matrix we get the accuracy of the model, more than 85%, pretty high, but we will try with the random forest algorithm to see if we get better results

#Training the new model.

```{r}
library(randomForest)
modelFit2 <- randomForest(classe ~. , data=training)
print(modelFit2)
```

#Evaluating the new model.
```{r}
predictions2 <- predict(modelFit2, testing, type = "class")
confusionMatrix(predictions2, testing$classe)
```
###We can see that applying random forest we get better accuracy and less missclassifications.